The all-seeing eye of resonant Auger electron spectroscopy: a study on aqueous KCl

### 2018-01-19 master.tex
theoretical XAS spectra + corrections to the text (jan 19)
Denis' corrections from Jan 18, references fixed, new theoretical XAS
spectrum

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]
  {The electronic structure of aqueous KCl revealed by X-ray absorption and Auger electron spectroscopy
\\  
  The all-seeing eye of Auger electron spectroscopy: a study on aqueous KCl}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{AES,XAS}
\keywords{Solvated ions, Auger spectroscopy, X-ray absorption spectroscopy}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]
  {The all-seeing eye of resonant Auger electron spectroscopy: a study on aqueous KCl}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{AES,XAS}
\keywords{Solvated ions, resonant Auger spectroscopy, X-ray absorption spectroscopy}

### 2018-01-19 methods.tex
theoretical XAS spectra + corrections to the text (jan 19)
Denis' corrections from Jan 18, references fixed, new theoretical XAS
spectrum

\section{Methods} \label{sec:methods}
\subsection{Experimental}
% Denis, please have a look at the experimental part and see if all the references have been correctly included

For the present experiment we used the newly operational microjet setup that was specifically designed for the HAXPES station of the GALAXIES beamline, in collaboration with the Microliquids Company. Details of the beamline are given in the reference [JPR] and details on the electron spectrometer are given in the reference \citep{ceolin15:022502,ceolin13:188}. Due to the limited amount of space available \kolya{in} front of the analyzer lens and due to the limited number of ports available on the main vacuum chamber, the challenge was to arrange all the elements composing the microjet setup in a very compact manner. A differentially-pumped tube in which the microjet assembly is inserted, is installed in front of the spectrometer lens and can be moved independently from the main vacuum chamber by means of a three axes motorized manipulator. Two holes of 2 mm diameter allow the photons to go in and out of the tube\kolya{. A third hole,} on which is mounted a skimmer having a 500$\mu$m diameter hole, allows the electrons created at the interaction point to go in the direction of the spectrometer lens. To ensure a proper vacuum in the differentially pumped tube, a {\color{red}5-way cross} is connected to it and holds a 1200 l/s turbo pump, a liquid nitrogen trap, a vacuum gage and the liquid/electrical feedthroughs. A rail, fixed at the bottom of the tube, is used to {\kolya slides} precisely an insert on which the different elements necessary for the injection, control, collection and visualization of the liquid are mounted.

The head of this insert is mostly composed by: a glass capillary fixed by a dedicated peek-piece, a catcher in CuBe with a 300$\mu$m hole, a piezo motors stage allowing a precise control of these last two parts, and a camera. The catcher is placed at a distance of about 5mm from the capillary and is permanently pumped in order to extract the liquid from the vacuum chamber. For the present experiment, a 0.5M KCl aqueous solution is injected in a capillary having a 30$\mu$m diameter, by a HPLC pump with a constant flux of 1.6 ml/min. The catcher temperature is controlled so that the liquid does not freeze before its extraction. Considering that the photon propagation axis, the spectrometer lens axis and the liquid microjet form an \kolya{orthogonal trihedron}, the piezo motorization allows moving the liquid microjet along the photon propagation and the lens axes during the experiment if necessary. The catcher can be moved by the piezo motors along the photon propagation axis only. A simple camera is used to control the liquid jet positioning as compared with the catcher hole. The alignment of the full head (particularly capillary plus catcher) as compared with the X-ray beam is performed by moving the whole system using the 3-axes motorized manipulator. The alignment of the setup is performed by measuring the O1s XPS peak intensity of a simple salt aqueous solution, and by optimizing the liquid phase vs gas phase ratio. The pressure in the main chamber was kept below the 10$^{-5}$ mbar range whereas it was kept at about 10$^{-4}$ mbar in the differentially pumped tube when the HPLC pump \kolya{was ON}. Our equipment is an updated version of the equipment used in the references {\color{red}WHAT r these refs [3-5]?}. The aqueous potassium chloride solution was prepared by mixing >99\% KCl salt with deionized water. Filtering and degazing procedures were systematically performed before injecting the solution into the microjet by the HPLC pump.

>>>>

\section{Methods} \label{sec:methods}
\subsection{Experimental}
% Denis, please have a look at the experimental part and see if all the references have been correctly included
For the present experiment we used the newly operational microjet setup that was specifically designed for the HAXPES station of the GALAXIES beamline \citep{ceolin13:188,rueff15:175}. A differentially-pumped tube in which the microjet head is inserted, is mounted on a 3-axes manipulator in front of the spectrometer lens. Two holes of 2\,mm diameter allow the photons to go in and out of the tube. At the end of the tube and in front of the lens, a 500\,$\mu$m diameter hole skimmer allows the electrons created at the interaction point to go in the direction of the spectrometer. The microjet head is mostly composed of a 30\,$\mu$m diameter vertical glass capillary facing a temperature-controlled catcher in CuBe having a 300\,$\mu$m hole, and a camera. Piezo motors allow their precise alignment relative to each other and to the photon beam. The catcher is placed at a distance of about 5\,mm from the capillary and is permanently pumped in order to extract the liquid. For the present experiment, a 0.5M KCl aqueous solution is injected in the capillary by a HPLC pump with a constant flux of 1.6\,ml/min. The alignment of the setup is performed on a salted aqueous solution by measuring the O1s XPS peak intensity and by optimizing the liquid phase vs gas phase ratio. The pressure in the main chamber is kept below the 10$^{−5}$\,mbar range whereas it is kept at about 10$^{−4}$\,mbar in the differentially-pumped tube when the HPLC pump is ON. Our equipment is an updated version of the equipment used in the reference \cite{faubel88:269}. The aqueous potassium chloride solution was prepared by mixing >99\% KCl salt with deionized water. Filtering and degazing procedures were systematically performed before injecting the solution. The spectrometer resolution of about 0.6 eV was achieved with the 500 eV pass energy and 0.5\,mm slits. The photon energy resolution achieved at 2.8\,keV and 3.6\,keV was about 0.3\,eV and 0.4\,eV, respectively

%
%For the present experiment we used the newly operational microjet setup that was specifically designed for the HAXPES station of the GALAXIES beamline, in collaboration with the Microliquids Company. Details of the beamline are given in the reference [JPR] and details on the electron spectrometer are given in the reference \citep{ceolin15:022502,ceolin13:188}. Due to the limited amount of space available \kolya{in} front of the analyzer lens and due to the limited number of ports available on the main vacuum chamber, the challenge was to arrange all the elements composing the microjet setup in a very compact manner. A differentially-pumped tube in which the microjet assembly is inserted, is installed in front of the spectrometer lens and can be moved independently from the main vacuum chamber by means of a three axes motorized manipulator. Two holes of 2 mm diameter allow the photons to go in and out of the tube\kolya{. A third hole,} on which is mounted a skimmer having a 500$\mu$m diameter hole, allows the electrons created at the interaction point to go in the direction of the spectrometer lens. To ensure a proper vacuum in the differentially pumped tube, a {\color{red}5-way cross} is connected to it and holds a 1200 l/s turbo pump, a liquid nitrogen trap, a vacuum gage and the liquid/electrical feedthroughs. A rail, fixed at the bottom of the tube, is used to {\kolya slides} precisely an insert on which the different elements necessary for the injection, control, collection and visualization of the liquid are mounted.
%
%
%The head of this insert is mostly composed by: a glass capillary fixed by a dedicated peek-piece, a catcher in CuBe with a 300$\mu$m hole, a piezo motors stage allowing a precise control of these last two parts, and a camera. The catcher is placed at a distance of about 5mm from the capillary and is permanently pumped in order to extract the liquid from the vacuum chamber. For the present experiment, a 0.5M KCl aqueous solution is injected in a capillary having a 30$\mu$m diameter, by a HPLC pump with a constant flux of 1.6 ml/min. The catcher temperature is controlled so that the liquid does not freeze before its extraction. Considering that the photon propagation axis, the spectrometer lens axis and the liquid microjet form an \kolya{orthogonal trihedron}, the piezo motorization allows moving the liquid microjet along the photon propagation and the lens axes during the experiment if necessary. The catcher can be moved by the piezo motors along the photon propagation axis only. A simple camera is used to control the liquid jet positioning as compared with the catcher hole. The alignment of the full head (particularly capillary plus catcher) as compared with the X-ray beam is performed by moving the whole system using the 3-axes motorized manipulator. The alignment of the setup is performed by measuring the O1s XPS peak intensity of a simple salt aqueous solution, and by optimizing the liquid phase vs gas phase ratio. The pressure in the main chamber was kept below the 10$^{-5}$ mbar range whereas it was kept at about 10$^{-4}$ mbar in the differentially pumped tube when the HPLC pump \kolya{was ON}. Our equipment is an updated version of the equipment used in the references {\color{red}WHAT r these refs [3-5]?}. The aqueous potassium chloride solution was prepared by mixing >99\% KCl salt with deionized water. Filtering and degazing procedures were systematically performed before injecting the solution into the microjet by the HPLC pump.

### 2018-01-19 results.tex
theoretical XAS spectra + corrections to the text (jan 19)
Denis' corrections from Jan 18, references fixed, new theoretical XAS
spectrum

\begin{figure}[h]
\centering
\includegraphics[scale=0.6]{figures/xas_spectra.eps}
\caption{
XAS spectra of the lowest K-shell transitions in the bare K$^{+}$ (lower left panel) and Cl$^{-}$ (lower right panel) ions and their 6-coordinated clusters (upper left panel, \ki(H$_2$O)$_6$, and upper right panel, \cli(H$_2$O)$_6$). The theoretical stick spectra were convolved with a Lorentzian profile of FWHM 0.74\,eV for \ki~and 0.62\,eV for \cli~in order to account for the lifetime broadening. The stick spectrum corresponds to the projections $|a_{nl}^{i}|^2$ of the SONOs corresponding to the core excited states of the 6-coordinated clusters on the basis of SONOs corresponding to the 1s$\,\rightarrow\,$3d, 1s$\,\rightarrow\,$4p, and 1s$\,\rightarrow\,$5p states in the bare K$^+$ and \cli~ions (Eq.\ \ref{eq:sono_proj}). The theoretical XAS spectra of \ki~were shifted to higher photon energies by 6.7\,eV, which corresponds to the difference between the computed and experimental core excitation energies of the 1s $\rightarrow$ 4p excitation in the bare ion taken from Ref.\ \citep{hertlein06:062715}. \tsveta{A DARKER SHADE OF GREY}}
\label{fg:xas_kcl}
\end{figure}

In order to rationalize the pre-edge region of the experimental XAS spectra and the differences in the AES spectra of \ki$_{\text{aq}}$ and \cli$_{\text{aq}}$, we computed the lowest core excited states of the bare \ki~and \cli~ions and their hexa-coordinated clusters. The theoretical XAS spectra are presented in Fig.\ \ref{fg:xas_kcl}. In the bare ions (lowermost panels on Fig.\ \ref{fg:xas_kcl}), the lowest energy peak corresponds to the dipole allowed 1s$\,\rightarrow\,$4p state. The next dipole allowed state, 1s$\,\rightarrow\,$5p, is located 4.3\,eV and 10.8\,eV higher in the cases of \ki~and \cli, respectively. Together with the two dipole allowed transitions, we show the dipole forbidden 1s$\,\rightarrow\,$3d states of the bare ions as blue crosses at photon energies \tsveta{3609.XX\,eV in the case of \ki~and 2831\,eV in the case of \cli, respectively}. It is noteworthy that the positions of the 1s$\,\rightarrow\,$4p and 1s$\,\rightarrow\,$3d states are inverted in \ki~and \cli. In the case of Cl$^{-}$ the 1s$\,\rightarrow\,$4p excitation has lower energy and the 1s$\,\rightarrow\,$3d excitation is close to the 1s$\,\rightarrow\,$5p state. On the contrary, in K$^{+}$ the 1s$\,\rightarrow\,$3d excitation has lower energy and lies below the 1s$\,\rightarrow\,$4p state. We note in passing that the intensity of the Cl$^{-}$(1s$\,\rightarrow\,$4p) state is lower than that of the \cli(1s$\,\rightarrow\,$5p) state contrary to what is observed in \ki. This difference can be explained with the lower electron density of the 4p compared to the 5p electron in the region close to the core hole which thus results in the lower oscillator strength of the 1s$\,\rightarrow\,$4p compared to the 1s$\,\rightarrow\,$5p transition in \cli~(see Fig.\ \ref{fg:rdens_ions}).

@@

In the hexa-coordinated cluster of \cli, the solvent molecules have little influence on the position and character of the first peak. Since there are no other ionic states close the the 1s$\,\rightarrow\,$4p state in the ion, this state preserves its character in the cluster and interacts with states of the nearest water molecules. We attribute the two dispersive features on the 2D map of \cli~to the resonant Auger decay of this core excited state. {\color{red}Since the photon energy step in the experimental spectrum coincides with the splitting between the two dispersive features, we assume that they originate from the decay of the same core excited state.}

To fully characterize the dispersive features on the experimental 2D maps, we also computed the lowest K$^{2+}$(2p$^{-2}$nl) and Cl$^{0}$(2p$^{-2}$nl) states of the bare ions corresponding to the lowest final spectator resonant Auger states. They are shown as bars in the upper panels on the experimental resonant Auger spectra on Figs.\ \ref{fg:2dmap_k} and \ref{fg:2dmap_cl}. In both cases of \ki~and \cli, we shifted the lowest 2p$^{-2}$4p states such that they coincide with the maxima of the dispersive features on the high kinetic energy part of the $^1$D main line. Out of all final states we show only the doublets. Since the initial core excited states are populated by photon excitation, only doublet states are populated in the resonant Auger decay. The 2p$^{-2}$nl states of \ki~and \cli~are substantially different and they reflect the fact that the 3d unoccupied orbitals of \ki~are lower than the 4p orbitals, which is the opposite of what is observed in \cli.

%As mentioned above, we assume that the island B is produced in the decay of the lowest lying core excited state of the hexa-hydrated \ki~cluster, which is of predominantly 1s$\,\rightarrow\,$4p character. 
As mentioned above, we attribute the island B to the decay of the lowest lying core excited state of the hexa-hydrated \ki~cluster, which is of predominantly 1s$\,\rightarrow\,$4p character. 
Supposing that this state undergoes mostly pure spectator resonant Auger decay, which is the case of the 1s$\,\rightarrow\,$4p state in the isoelectronic Ar atom \citep{ceolin15:022502}, then the lowest states of 2p$^{-2}$4p character located between 2969 and 2970\,eV are populated in this case. As can be seen from the Auger electron spectrum at $h\nu = $3610.7\,eV (upper panel of Fig.\ \ref{fg:2dmap_k}), the lowest 2p$^{-2}$4p states of \ki~are separated by 5 and 13\,eV from the lowest 2p$^{-2}$3d states. Thus, the lower-kinetic energy group of 2p$^{-2}$3d states almost coincides with the position of the island C. Consequently, we attribute this dispersive feature as originating from the resonant Auger decay of the dipole forbidden 1s$\,\rightarrow\,$3d state of the 6-coordinated \ki~cluster to the group of 2p$^{-2}$3d states lying around 2975\,eV. The splitting between the 2p$^{-2}$4p and 2p$^{-2}$3d states in our calculation is smaller than the splitting between the islands B and C. This difference may be due to the fact that we do not account for the effect of the solvent molecules in our calculation. There is a group of 2p$^{-2}$3d states at kinetic energies of 2982\,eV. Since no additional experimental features are observed, we conclude that these states are not populated.

>>>>

\begin{figure}[h]
\centering
\includegraphics[scale=0.6]{figures/xas_spectra.eps}
\caption{
XAS spectra of the lowest K-shell transitions in the bare K$^{+}$ (lower left panel) and Cl$^{-}$ (lower right panel) ions and their 6-coordinated clusters (upper left panel, \ki(H$_2$O)$_6$, and upper right panel, \cli(H$_2$O)$_6$). The theoretical stick spectra were convolved with a Lorentzian profile of FWHM 0.74\,eV for \ki~and 0.62\,eV for \cli~(dashed line) and a Voigt profile to account for the lifetime broadening and the experimental resolution. The stick spectrum corresponds to the projections $|a_{nl}^{i}|^2$ of the SONOs corresponding to the core excited states of the 6-coordinated clusters on the basis of SONOs corresponding to the 1s$\,\rightarrow\,$3d, 1s$\,\rightarrow\,$4p, and 1s$\,\rightarrow\,$5p states in the bare K$^+$ and \cli~ions (Eq.\ \ref{eq:sono_proj}). The theoretical XAS spectra of both \ki~and \cli~were shifted to higher photon energies such that the excitation energies lowest core excited states correspond to the experimentally determined energies -- 3610.7\,eV in the case of \ki, and 2825.2\,eV in the case of \cli. The experimental ionization thresholds are depicted as grey boxes starting at photon energies of 3611.9\,eV (\ki$_{\text{aq}}$) and 2825.4\,eV (\cli$_{\text{aq}}$).}
\label{fg:xas_kcl}
\end{figure}

In order to rationalize the pre-edge region of the experimental XAS spectra and the differences in the AES spectra of \ki$_{\text{aq}}$ and \cli$_{\text{aq}}$, we computed the lowest core excited states of the bare \ki~and \cli~ions and their hexa-coordinated clusters. The theoretical XAS spectra are presented in Fig.\ \ref{fg:xas_kcl}. In the bare ions (lowermost panels on Fig.\ \ref{fg:xas_kcl}), the lowest energy peak corresponds to the dipole allowed 1s$\,\rightarrow\,$4p state. The next dipole allowed state, 1s$\,\rightarrow\,$5p, is located 4.3\,eV and 10.8\,eV higher in the cases of \ki~and \cli, respectively. Together with the two dipole allowed transitions, we show the dipole forbidden 1s$\,\rightarrow\,$3d states of the bare ions as blue crosses at photon energies 3611.17\,eV in the case of \ki~and 2831.68\,eV in the case of \cli, respectively. It is noteworthy that the positions of the 1s$\,\rightarrow\,$4p and 1s$\,\rightarrow\,$3d states are inverted in \ki~and \cli. In the case of Cl$^{-}$ the 1s$\,\rightarrow\,$4p excitation has lower energy and the 1s$\,\rightarrow\,$3d excitation is close to the 1s$\,\rightarrow\,$5p state. On the contrary, in K$^{+}$ the 1s$\,\rightarrow\,$3d excitation has lower energy and lies below the 1s$\,\rightarrow\,$4p state. We note in passing that the intensity of the Cl$^{-}$(1s$\,\rightarrow\,$4p) state is lower than that of the \cli(1s$\,\rightarrow\,$5p) state contrary to what is observed in \ki. This difference can be explained with the lower electron density of the 4p compared to the 5p electron in the region close to the core hole which thus results in the lower oscillator strength of the 1s$\,\rightarrow\,$4p compared to the 1s$\,\rightarrow\,$5p transition in \cli~(see Fig.\ \ref{fg:rdens_ions}).

@@

In the hexa-coordinated cluster of \cli, the solvent molecules have little influence on the position and character of the first peak. Since there are no other ionic states close the the 1s$\,\rightarrow\,$4p state in the ion, this state preserves its character in the cluster and interacts with states of the nearest water molecules. We attribute the two dispersive features on the 2D map of \cli~to the resonant Auger decay of this core excited state. {\color{red}%Since the photon energy step in the experimental spectrum coincides with the splitting between the two dispersive features, we assume that they originate from the decay of the same core excited state.
We assume that the two dispersive features originate from the decay of the same core excited state, since the photon energy resolution at 2.8\,keV is 0.3\,eV and thus of the same magnitude as the difference between the core excited state and the ionization threshold.
}

To fully characterize the dispersive features on the experimental 2D maps, we also computed the lowest K$^{2+}$(2p$^{-2}$nl) and Cl$^{0}$(2p$^{-2}$nl) states of the bare ions corresponding to the lowest final spectator resonant Auger states. They are shown as bars in the upper panels on the experimental resonant Auger spectra on Figs.\ \ref{fg:2dmap_k} and \ref{fg:2dmap_cl}. In both cases of \ki~and \cli, we shifted the lowest 2p$^{-2}$4p states such that they coincide with the maxima of the dispersive features on the high kinetic energy part of the $^1$D main line. Out of all final states we show only the doublets. Since the initial core excited states are populated by photon excitation, only doublet states are efficiently populated in the resonant Auger decay. The 2p$^{-2}$nl states of \ki~and \cli~are substantially different and they reflect the fact that the 3d unoccupied orbitals of \ki~are lower than the 4p orbitals, which is the opposite of what is observed in \cli.

%As mentioned above, we assume that the island B is produced in the decay of the lowest lying core excited state of the hexa-hydrated \ki~cluster, which is of predominantly 1s$\,\rightarrow\,$4p character. 
As mentioned above, we attribute the island B to the decay of the lowest lying core excited state of the hexa-hydrated \ki~cluster, which is of predominantly 1s$\,\rightarrow\,$4p character. Supposing that this state undergoes mostly pure spectator resonant Auger decay, which is the case of the 1s$\,\rightarrow\,$4p state in the isoelectronic Ar atom \citep{ceolin15:022502}, then the lowest states of 2p$^{-2}$4p character located between 2969 and 2970\,eV are populated in this case. As can be seen from the Auger electron spectrum at $h\nu = $3610.7\,eV (upper panel of Fig.\ \ref{fg:2dmap_k}), the lowest 2p$^{-2}$4p states of \ki~at electron energy of $\sim$2970\,eV are separated by $\sim$5 and $\sim$13\,eV from the two groups of 2p$^{-2}$3d states located at higher electron energies. Thus, the group of 2p$^{-2}$3d states at $\sim$2975\,eV coincides with the position of island C. Consequently, we attribute this dispersive feature as originating from the resonant Auger decay of the dipole forbidden 1s$\,\rightarrow\,$3d state of the 6-coordinated \ki~cluster to the group of 2p$^{-2}$3d states lying around 2975\,eV. The splitting between the 2p$^{-2}$4p and 2p$^{-2}$3d states in our calculation is smaller than the splitting between the islands B and C. This difference may be due to the fact that we do not account for the effect of the solvent molecules in our calculation. There is a group of 2p$^{-2}$3d states at kinetic energies of 2982\,eV. Since no additional experimental features are observed, we conclude that these states are not populated.
